Equivalence Testing and the Second Generation \emph{P}-Value

### 2019-04-11 manuscript.Rmd
Correcting the misinterpretation of the p-value in correlation tests

The reason that the equivalence test and SGPV no longer overlap is also because of asymmetric confidence intervals. If the observed correlation falls exactly on the equivalence bound the *p*-value for the equivalence test indicates that the probability of observing the observed or more extreme data, assuming the equivalence bound is the true effect size, is 50%. In other words, if the true effect size is the same as the equivalence bound, it is equally likely to find an effect more extreme than the equivalence bound, as it is to observe an effect that is less extreme than the equivalence bound. However, as can be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation *p*-values associated with the observed correlations at r = -0.45 and r = 0.45 are larger than 50%. Because the confidence intervals are asymmetric around the observed effect size of 0.45 (ranging from `r round(res$LL_CI_TTEST,2)` to `r round(res$UL_CI_TTEST,2)`) according to @blume_second-generation_2018 `r round(sgpv*100,3)`% of the data-supported hypotheses are null hypotheses, and therefore `r round(sgpv,3)*100`% of the data-supported hypotheses are compatible with the null premise.

@@

This example illustrates the difference between a proportion and a probability. There is always a 50% probability of observing a correlation smaller or larger than the true correlation, but the SGPV for this situation depends on how far away the observed correlation is from 0. The further away from 0, the larger the SGPV when the observed mean falls on the equivalence bound. The SGPV is the proportion of values in a 95% confidence interval that overlap with the equivalence range, but not the probability that these values will be observed. In the most extreme case (i.e., a sample size of 4, and equivalence bounds set to r = -0.99 and 0.99, with an observed correlation of 0.99) `r round(sgpv,3)*100`% of the confidence interval overlaps with the equivalence range, even though in the long run only 50% of the correlations observed in the future will fall in this range. It should be noted that in larger sample sizes the SGPV is closer to 0.5 whenever the observed correlation falls on the equivalence bound, but this extreme example nevertheless clearly illustrates the difference between question the SGPV answers, and the question a *p*-value answers. The conclusion of this section on asymmetric confidence intervals is that a SGPV of 1 or 0 can still be interpreted as a *p* < 0.025 or *p* > 0.975 in an equivalence test, since the SGPV and *p*-value for the TOST procedure are always directly related at these values. Although @blume_second-generation_2018 state that "the degree of overlap conveys how compatible the data are with the null premise" this definition of what the SGPV provides does not hold for asymmetric confidence intervals. Although a SGPV of 1 or 0 can be directly interpreted, a SGPV between 0 and 1 is not interpretable as 'compatibility with the null hypothesis'. Indeed, Blume and colleagues write in the supplemental material that "The magnitude of an inconclusive second-generation *p*-value can vary slightly when the effect size scale is transformed. However definitive findings, i.e. a *p*-value of 0 or 1 are *not* affected by the scale changes."

>>>>

The reason that the equivalence test and SGPV no longer overlap is also because of asymmetric confidence intervals.

If the observed correlation falls exactly on the equivalence bound the *p*-value for the equivalence test is 50%. In correlation test, the *p* value is computed based on a transformed correlation score, computed as follows:

$$
z = \frac{\frac{\log(\frac{1 + r}{ 1 - r})}{2} - \frac{\log(\frac{1 + \rho}{ 1 - \rho})}{2}}{\sqrt{\frac{1}{n-3}}}
$$

Because the z distribution is symmetric, the probability of observing the observed z score or more extreme score, assuming the equivalence bound is the true effect size, is 50%. However, it does not mean that there is always a 50% probability of observing a correlation (r) smaller or larger than the true correlation: while the z distribution is symmetrically distributed around the observed z score, the r distribution is not.

As can be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation *p*-values associated with the observed correlations at r = -0.45 and r = 0.45 are larger than 50%. Because the confidence intervals are asymmetric around the observed effect size of 0.45 (ranging from `r round(res$LL_CI_TTEST,2)` to `r round(res$UL_CI_TTEST,2)`) according to @blume_second-generation_2018 `r round(sgpv*100,3)`% of the data-supported hypotheses are null hypotheses, and therefore `r round(sgpv,3)*100`% of the data-supported hypotheses are compatible with the null premise. The further away from 0, the larger the SGPV when the observed mean falls on the equivalence bound. The SGPV is the proportion of values in a 95% confidence interval that overlap with the equivalence range, but not the probability that these values will be observed. In the most extreme case (i.e., a sample size of 4, and equivalence bounds set to r = -0.99 and 0.99, with an observed correlation of 0.99) `r round(sgpv,3)*100`% of the confidence interval overlaps with the equivalence range, even though in the long run only 50% of the correlations observed in the future will fall in this range.

It should be noted that in larger sample sizes the SGPV is closer to 0.5 whenever the observed correlation falls on the equivalence bound, and in the same way of thinking, the probability of observing a correlation smaller or larger than the true correlation is also closer to 0.5.

The conclusion of this section on asymmetric confidence intervals is that a SGPV of 1 or 0 can still be interpreted as a *p* < 0.025 or *p* > 0.975 in an equivalence test, since the SGPV and *p*-value for the TOST procedure are always directly related at these values. Although @blume_second-generation_2018 state that "the degree of overlap conveys how compatible the data are with the null premise" this definition of what the SGPV provides does not hold for asymmetric confidence intervals. Although a SGPV of 1 or 0 can be directly interpreted, a SGPV between 0 and 1 is not interpretable as 'compatibility with the null hypothesis'. Indeed, Blume and colleagues write in the supplemental material that "The magnitude of an inconclusive second-generation *p*-value can vary slightly when the effect size scale is transformed. However definitive findings, i.e. a *p*-value of 0 or 1 are *not* affected by the scale changes."

@@

### 2019-04-11 manuscript.tex
Correcting the misinterpretation of the p-value in correlation tests

\documentclass[floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Equivalence Testing and the Second Generation P-Value},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Equivalence Testing and the Second Generation \emph{P}-Value}

  \shorttitle{TOST vs.~SGPV}

  \author{Daniël Lakens\textsuperscript{1}~\& Marie Delacre\textsuperscript{2}}

  \def\affdep{{"", ""}}%
  \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Eindhoven University of Technology, Eindhoven, The Netherlands\\
          \textsuperscript{2} Service of Analysis of the Data, Université Libre de Bruxelles, Belgium  }

  \authornote{
    \newcounter{author}
    All code associated with this article, including the reproducible
    manuscript, is available from
    \url{https://github.com/Lakens/TOST_vs_SGPV}. This work was supported by
    the Netherlands Organization for Scientific Research (NWO) VIDI grant
    452-17-013.

                      Correspondence concerning this article should be addressed to Daniël Lakens, Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands. E-mail: \href{mailto:D.Lakens@tue.nl}{\nolinkurl{D.Lakens@tue.nl}}
                          }

  \abstract{To move beyond the limitations of null-hypothesis tests, statistical
approaches have been developed where the observed data is compared
against a range of values that are equivalent to the absence of a
meaningful effect. Specifying a range of values around zero allows
researchers to statistically reject the presence of effects large enough
to matter, and prevents practically insignificant effects from being
interpreted as a statistically significant difference. We compare the
behavior of the recently proposed second generation \emph{p}-value
(Blume, D'Agostino McGowan, Dupont, \& Greevy, 2018) with the more
established Two One-Sided Tests (TOST) equivalence testing procedure
(Schuirmann, 1987). We show that the two approaches yield almost
identical results under optimal conditions. Under suboptimal conditions
(e.g., when the confidence interval is wider than the equivalence range,
or when confidence intervals are asymmetric) the second generation
\emph{p}-value becomes difficult to interpret as a descriptive
statistic. The second generation \emph{p}-value is interpretable in a
dichotomous manner (i.e., when the SGPV equals 0 or 1 because the
confidence intervals lies completely within or outside of the
equivalence range), but this dichotomous interpretation does not require
calculations. We conclude that equivalence tests yield more consistent
\emph{p}-values, distinguish between datasets that yield the same second
generation \emph{p}-value, and allow for easier control of Type I and
Type II error rates.}
  \keywords{equivalence testing, second generation \emph{p}-values, hypothesis
testing, TOST, statistical inference \\

@@

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

@@

\section{\texorpdfstring{The relationship between \emph{p}-values from
TOST and SGPV when confidence intervals are
symmetrical}{The relationship between p-values from TOST and SGPV when confidence intervals are symmetrical}}\label{the-relationship-between-p-values-from-tost-and-sgpv-when-confidence-intervals-are-symmetrical}

@@

\subsection{SGPV as a uniform measure of
overlap}\label{sgpv-as-a-uniform-measure-of-overlap}

@@

\subsection{When are the SGPV and Equivalence Test
Unrelated?}\label{when-are-the-sgpv-and-equivalence-test-unrelated}

@@

\subsection{The relation between equivalence tests and SGPV when
confidence intervals are not
symmetrical}\label{the-relation-between-equivalence-tests-and-sgpv-when-confidence-intervals-are-not-symmetrical}

@@

The reason that the equivalence test and SGPV no longer overlap is also
because of asymmetric confidence intervals. If the observed correlation
falls exactly on the equivalence bound the \emph{p}-value for the
equivalence test indicates that the probability of observing the
observed or more extreme data, assuming the equivalence bound is the
true effect size, is 50\%. In other words, if the true effect size is
the same as the equivalence bound, it is equally likely to find an
effect more extreme than the equivalence bound, as it is to observe an
effect that is less extreme than the equivalence bound. However, as can
be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation
\emph{p}-values associated with the observed correlations at r = -0.45
and r = 0.45 are larger than 50\%. Because the confidence intervals are
asymmetric around the observed effect size of 0.45 (ranging from 0.11 to
0.70) according to Blume et al. (2018) 58.11\% of the data-supported
hypotheses are null hypotheses, and therefore 58.10\% of the
data-supported hypotheses are compatible with the null premise.

@@

This example illustrates the difference between a proportion and a
probability. There is always a 50\% probability of observing a
correlation smaller or larger than the true correlation, but the SGPV
for this situation depends on how far away the observed correlation is
from 0. The further away from 0, the larger the SGPV when the observed
mean falls on the equivalence bound. The SGPV is the proportion of
values in a 95\% confidence interval that overlap with the equivalence
range, but not the probability that these values will be observed. In
the most extreme case (i.e., a sample size of 4, and equivalence bounds
set to r = -0.99 and 0.99, with an observed correlation of 0.99) 58.10\%
of the confidence interval overlaps with the equivalence range, even
though in the long run only 50\% of the correlations observed in the
future will fall in this range. It should be noted that in larger sample
sizes the SGPV is closer to 0.5 whenever the observed correlation falls
on the equivalence bound, but this extreme example nevertheless clearly
illustrates the difference between question the SGPV answers, and the
question a \emph{p}-value answers. The conclusion of this section on
asymmetric confidence intervals is that a SGPV of 1 or 0 can still be
interpreted as a \emph{p} \textless{} 0.025 or \emph{p} \textgreater{}
0.975 in an equivalence test, since the SGPV and \emph{p}-value for the
TOST procedure are always directly related at these values. Although
Blume et al. (2018) state that \enquote{the degree of overlap conveys
how compatible the data are with the null premise} this definition of
what the SGPV provides does not hold for asymmetric confidence
intervals. Although a SGPV of 1 or 0 can be directly interpreted, a SGPV
between 0 and 1 is not interpretable as \enquote{compatibility with the
null hypothesis}. Indeed, Blume and colleagues write in the supplemental
material that \enquote{The magnitude of an inconclusive
second-generation \emph{p}-value can vary slightly when the effect size
scale is transformed. However definitive findings, i.e.~a \emph{p}-value
of 0 or 1 are \emph{not} affected by the scale changes.}

\subsection{What are the Relative Strengths and Weaknesses of
Equivalence Testing and the
SGPV?}\label{what-are-the-relative-strengths-and-weaknesses-of-equivalence-testing-and-the-sgpv}

@@

\section{Conclusion}\label{conclusion}

@@

\section{References}\label{references}

@@

\hypertarget{refs}{}
\hypertarget{ref-american_psychological_association_publication_2010}{}
American Psychological Association (Ed.). (2010). \emph{Publication
manual of the American Psychological Association} (6th ed.). Washington,
DC: American Psychological Association.

\hypertarget{ref-blume_second-generation_2018}{}
Blume, J. D., D'Agostino McGowan, L., Dupont, W. D., \& Greevy, R. A.
(2018). Second-generation p-values: Improved rigor, reproducibility, \&
transparency in statistical analyses. \emph{PLOS ONE}, \emph{13}(3),
e0188299.
doi:\href{https://doi.org/10.1371/journal.pone.0188299}{10.1371/journal.pone.0188299}

\hypertarget{ref-hauck_new_1984}{}
Hauck, D. W. W., \& Anderson, S. (1984). A new statistical procedure for
testing equivalence in two-group comparative bioavailability trials.
\emph{Journal of Pharmacokinetics and Biopharmaceutics}, \emph{12}(1),
83--91.
doi:\href{https://doi.org/10.1007/BF01063612}{10.1007/BF01063612}

\hypertarget{ref-kruschke_rejecting_2018}{}
Kruschke, J. K. (2018). Rejecting or Accepting Parameter Values in
Bayesian Estimation. \emph{Advances in Methods and Practices in
Psychological Science}, 2515245918771304.
doi:\href{https://doi.org/10.1177/2515245918771304}{10.1177/2515245918771304}

\hypertarget{ref-lakens_equivalence_2017}{}
Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests,
Correlations, and Meta-Analyses. \emph{Social Psychological and
Personality Science}, \emph{8}(4), 355--362.
doi:\href{https://doi.org/10.1177/1948550617697177}{10.1177/1948550617697177}

\hypertarget{ref-lakens_equivalence_2018}{}
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence Testing
for Psychological Research: A Tutorial. \emph{Advances in Methods and
Practices in Psychological Science}, 2515245918770963.
doi:\href{https://doi.org/10.1177/2515245918770963}{10.1177/2515245918770963}

\hypertarget{ref-meyners_equivalence_2012}{}
Meyners, M. (2012). Equivalence tests review. \emph{Food Quality and
Preference}, \emph{26}(2), 231--245.
doi:\href{https://doi.org/10.1016/j.foodqual.2012.05.003}{10.1016/j.foodqual.2012.05.003}

\hypertarget{ref-murphy_statistical_2014}{}
Murphy, K. R., Myors, B., \& Wolach, A. H. (2014). \emph{Statistical
power analysis: A simple and general model for traditional and modern
hypothesis tests} (Fourth edition.). New York: Routledge, Taylor \&
Francis Group.

\hypertarget{ref-quertemont_how_2011}{}
Quertemont, E. (2011). How to Statistically Show the Absence of an
Effect. \emph{Psychologica Belgica}, \emph{51}(2), 109--127.
doi:\href{https://doi.org/10.5334/pb-51-2-109}{10.5334/pb-51-2-109}

\hypertarget{ref-rogers_using_1993}{}
Rogers, J. L., Howard, K. I., \& Vessey, J. T. (1993). Using
significance tests to evaluate equivalence between two experimental
groups. \emph{Psychological Bulletin}, \emph{113}(3), 553--565.
doi:\href{https://doi.org/http://dx.doi.org/10.1037/0033-2909.113.3.553}{http://dx.doi.org/10.1037/0033-2909.113.3.553}

\hypertarget{ref-schuirmann_comparison_1987}{}
Schuirmann, D. J. (1987). A comparison of the two one-sided tests
procedure and the power approach for assessing the equivalence of
average bioavailability. \emph{Journal of Pharmacokinetics and
Biopharmaceutics}, \emph{15}(6), 657--680.

\hypertarget{ref-serlin_rationality_1985}{}
Serlin, R. C., \& Lapsley, D. K. (1985). Rationality in psychological
research: The good-enough principle.

\hypertarget{ref-spiegelhalter_bayesian_1994}{}
Spiegelhalter, D. J., Freedman, L. S., \& Parmar, M. K. (1994). Bayesian
approaches to randomized trials. \emph{Journal of the Royal Statistical
Society. Series A (Statistics in Society)}, 357--416.
doi:\href{https://doi.org/10.2307/2983527}{10.2307/2983527}

\hypertarget{ref-wellek_testing_2010}{}
Wellek, S. (2010). \emph{Testing statistical hypotheses of equivalence
and noninferiority} (2nd ed.). Boca Raton: CRC Press.

\hypertarget{ref-westlake_use_1972}{}
Westlake, W. J. (1972). Use of Confidence Intervals in Analysis of
Comparative Bioavailability Trials. \emph{Journal of Pharmaceutical
Sciences}, \emph{61}(8), 1340--1341.
doi:\href{https://doi.org/10.1002/JPS.2600610845}{10.1002/JPS.2600610845}

>>>>

\documentclass[,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Equivalence Testing and the Second Generation P-Value},
            pdfauthor={Daniël Lakens~\& Marie Delacre},
            pdfkeywords={equivalence testing, second generation \emph{p}-values, hypothesis testing, TOST, statistical inference},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Equivalence Testing and the Second Generation \emph{P}-Value}
    \author{Daniël Lakens\textsuperscript{1}~\& Marie Delacre\textsuperscript{2}}
    \date{}

\shorttitle{TOST vs. SGPV}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Eindhoven University of Technology, Eindhoven, The Netherlands\\\textsuperscript{2} Service of Analysis of the Data, Université Libre de Bruxelles, Belgium}
\keywords{equivalence testing, second generation *p*-values, hypothesis testing, TOST, statistical inference\newline\indent Word count: }
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\authornote{All code associated with this article, including the reproducible manuscript, is available from \url{https://github.com/Lakens/TOST_vs_SGPV}. This work was supported by the Netherlands Organization for Scientific Research (NWO) VIDI grant 452-17-013.

Correspondence concerning this article should be addressed to Daniël Lakens, Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands. E-mail: \href{mailto:D.Lakens@tue.nl}{\nolinkurl{D.Lakens@tue.nl}}}

\abstract{
To move beyond the limitations of null-hypothesis tests, statistical approaches have been developed where the observed data is compared against a range of values that are equivalent to the absence of a meaningful effect. Specifying a range of values around zero allows researchers to statistically reject the presence of effects large enough to matter, and prevents practically insignificant effects from being interpreted as a statistically significant difference. We compare the behavior of the recently proposed second generation \emph{p}-value (Blume, D'Agostino McGowan, Dupont, \& Greevy, 2018) with the more established Two One-Sided Tests (TOST) equivalence testing procedure (Schuirmann, 1987). We show that the two approaches yield almost identical results under optimal conditions. Under suboptimal conditions (e.g., when the confidence interval is wider than the equivalence range, or when confidence intervals are asymmetric) the second generation \emph{p}-value becomes difficult to interpret as a descriptive statistic. The second generation \emph{p}-value is interpretable in a dichotomous manner (i.e., when the SGPV equals 0 or 1 because the confidence intervals lies completely within or outside of the equivalence range), but this dichotomous interpretation does not require calculations. We conclude that equivalence tests yield more consistent \emph{p}-values, distinguish between datasets that yield the same second generation \emph{p}-value, and allow for easier control of Type I and Type II error rates.

@@

\begin{document}
\maketitle

@@

\#The relationship between \emph{p}-values from TOST and SGPV when confidence intervals are symmetrical

@@

\#\#SGPV as a uniform measure of overlap

@@

\#\#When are the SGPV and Equivalence Test Unrelated?

@@

\#\#The relation between equivalence tests and SGPV when confidence intervals are not symmetrical

@@

The reason that the equivalence test and SGPV no longer overlap is also because of asymmetric confidence intervals.

If the observed correlation falls exactly on the equivalence bound the \emph{p}-value for the equivalence test is 50\%. In correlation test, the \emph{p} value is computed based on a transformed correlation score, computd as follows:

\[
z = \frac{\frac{\log(\frac{1 + r}{ 1 - r})}{2} - \frac{\log(\frac{1 + \rho}{ 1 - \rho})}{2}}{\sqrt{\frac{1}{n-3}}}
\]

Because the z distribution is symmetric, the probability of observing the observed z score or more extreme score, assuming the equivalence bound is the true effect size, is 50\%. However, it does not mean that there is always a 50\% probability of observing a correlation (r) smaller or larger than the true correlation: while the z distribution is symmetrically distributed around the observed z score, the r distribution is not.

As can be seen in Figure \ref{fig:TOSTSGPV12}, the two second generation \emph{p}-values associated with the observed correlations at r = -0.45 and r = 0.45 are larger than 50\%. Because the confidence intervals are asymmetric around the observed effect size of 0.45 (ranging from 0.11 to 0.70) according to Blume et al. (2018) 58.11\% of the data-supported hypotheses are null hypotheses, and therefore 58.10\% of the data-supported hypotheses are compatible with the null premise. The further away from 0, the larger the SGPV when the observed mean falls on the equivalence bound. The SGPV is the proportion of values in a 95\% confidence interval that overlap with the equivalence range, but not the probability that these values will be observed. In the most extreme case (i.e., a sample size of 4, and equivalence bounds set to r = -0.99 and 0.99, with an observed correlation of 0.99) 58.10\% of the confidence interval overlaps with the equivalence range, even though in the long run only 50\% of the correlations observed in the future will fall in this range.

It should be noted that in larger sample sizes the SGPV is closer to 0.5 whenever the observed correlation falls on the equivalence bound, but this extreme example nevertheless clearly illustrates the difference between question the SGPV answers, and the question a \emph{p}-value answers. The conclusion of this section on asymmetric confidence intervals is that a SGPV of 1 or 0 can still be interpreted as a \emph{p} \textless{} 0.025 or \emph{p} \textgreater{} 0.975 in an equivalence test, since the SGPV and \emph{p}-value for the TOST procedure are always directly related at these values. Although Blume et al. (2018) state that \enquote{the degree of overlap conveys how compatible the data are with the null premise} this definition of what the SGPV provides does not hold for asymmetric confidence intervals. Although a SGPV of 1 or 0 can be directly interpreted, a SGPV between 0 and 1 is not interpretable as \enquote{compatibility with the null hypothesis}. Indeed, Blume and colleagues write in the supplemental material that \enquote{The magnitude of an inconclusive second-generation \emph{p}-value can vary slightly when the effect size scale is transformed. However definitive findings, i.e.~a \emph{p}-value of 0 or 1 are \emph{not} affected by the scale changes.}

@@

\#\#What are the Relative Strengths and Weaknesses of Equivalence Testing and the SGPV?

@@

\#Conclusion

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-american_psychological_association_publication_2010}{}%
American Psychological Association (Ed.). (2010). \emph{Publication manual of the American Psychological Association} (6th ed.). Washington, DC: American Psychological Association.

\leavevmode\hypertarget{ref-blume_second-generation_2018}{}%
Blume, J. D., D'Agostino McGowan, L., Dupont, W. D., \& Greevy, R. A. (2018). Second-generation p-values: Improved rigor, reproducibility, \& transparency in statistical analyses. \emph{PLOS ONE}, \emph{13}(3), e0188299. doi:\href{https://doi.org/10.1371/journal.pone.0188299}{10.1371/journal.pone.0188299}

\leavevmode\hypertarget{ref-hauck_new_1984}{}%
Hauck, D. W. W., \& Anderson, S. (1984). A new statistical procedure for testing equivalence in two-group comparative bioavailability trials. \emph{Journal of Pharmacokinetics and Biopharmaceutics}, \emph{12}(1), 83--91. doi:\href{https://doi.org/10.1007/BF01063612}{10.1007/BF01063612}

\leavevmode\hypertarget{ref-kruschke_rejecting_2018}{}%
Kruschke, J. K. (2018). Rejecting or Accepting Parameter Values in Bayesian Estimation. \emph{Advances in Methods and Practices in Psychological Science}, 2515245918771304. doi:\href{https://doi.org/10.1177/2515245918771304}{10.1177/2515245918771304}

\leavevmode\hypertarget{ref-lakens_equivalence_2017}{}%
Lakens, D. (2017). Equivalence Tests: A Practical Primer for t Tests, Correlations, and Meta-Analyses. \emph{Social Psychological and Personality Science}, \emph{8}(4), 355--362. doi:\href{https://doi.org/10.1177/1948550617697177}{10.1177/1948550617697177}

\leavevmode\hypertarget{ref-lakens_equivalence_2018}{}%
Lakens, D., Scheel, A. M., \& Isager, P. M. (2018). Equivalence Testing for Psychological Research: A Tutorial. \emph{Advances in Methods and Practices in Psychological Science}, 2515245918770963. doi:\href{https://doi.org/10.1177/2515245918770963}{10.1177/2515245918770963}

\leavevmode\hypertarget{ref-meyners_equivalence_2012}{}%
Meyners, M. (2012). Equivalence tests review. \emph{Food Quality and Preference}, \emph{26}(2), 231--245. doi:\href{https://doi.org/10.1016/j.foodqual.2012.05.003}{10.1016/j.foodqual.2012.05.003}

\leavevmode\hypertarget{ref-murphy_statistical_2014}{}%
Murphy, K. R., Myors, B., \& Wolach, A. H. (2014). \emph{Statistical power analysis: A simple and general model for traditional and modern hypothesis tests} (Fourth edition.). New York: Routledge, Taylor \& Francis Group.

\leavevmode\hypertarget{ref-quertemont_how_2011}{}%
Quertemont, E. (2011). How to Statistically Show the Absence of an Effect. \emph{Psychologica Belgica}, \emph{51}(2), 109--127. doi:\href{https://doi.org/10.5334/pb-51-2-109}{10.5334/pb-51-2-109}

\leavevmode\hypertarget{ref-rogers_using_1993}{}%
Rogers, J. L., Howard, K. I., \& Vessey, J. T. (1993). Using significance tests to evaluate equivalence between two experimental groups. \emph{Psychological Bulletin}, \emph{113}(3), 553--565. doi:\href{https://doi.org/http://dx.doi.org/10.1037/0033-2909.113.3.553}{http://dx.doi.org/10.1037/0033-2909.113.3.553}

\leavevmode\hypertarget{ref-schuirmann_comparison_1987}{}%
Schuirmann, D. J. (1987). A comparison of the two one-sided tests procedure and the power approach for assessing the equivalence of average bioavailability. \emph{Journal of Pharmacokinetics and Biopharmaceutics}, \emph{15}(6), 657--680.

\leavevmode\hypertarget{ref-serlin_rationality_1985}{}%
Serlin, R. C., \& Lapsley, D. K. (1985). Rationality in psychological research: The good-enough principle.

\leavevmode\hypertarget{ref-spiegelhalter_bayesian_1994}{}%
Spiegelhalter, D. J., Freedman, L. S., \& Parmar, M. K. (1994). Bayesian approaches to randomized trials. \emph{Journal of the Royal Statistical Society. Series A (Statistics in Society)}, 357--416. doi:\href{https://doi.org/10.2307/2983527}{10.2307/2983527}

\leavevmode\hypertarget{ref-wellek_testing_2010}{}%
Wellek, S. (2010). \emph{Testing statistical hypotheses of equivalence and noninferiority} (2nd ed.). Boca Raton: CRC Press.

\leavevmode\hypertarget{ref-westlake_use_1972}{}%
Westlake, W. J. (1972). Use of Confidence Intervals in Analysis of Comparative Bioavailability Trials. \emph{Journal of Pharmaceutical Sciences}, \emph{61}(8), 1340--1341. doi:\href{https://doi.org/10.1002/JPS.2600610845}{10.1002/JPS.2600610845}
